Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-20 manuscript.tex
final submission edits

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
%\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning of the slow dynamical modes, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Hinckley2014Coarse-grainedEffects,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of out-of-register shifted states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomena including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSimulations, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in good agreement with observed experimental trends.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and slithering mechanisms.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima were estimated using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but note that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

@@

In Fig.~\ref{fig:relaxation-comparison}a we observe an exponential increase of $k_d^\mathrm{slow}$ with temperature, as expected from the large enthalpic barrier to duplex dissociation.\cite{Craig1971RelaxationOligonucleotides,Porschke1971CooperativeTransition,Williams1989LaserDGCATGC} Under the time and temperature calibration corrections, we see generally very good agreement between the computational and experimental curves. The computational rates do tend to systematically under-predict the experimental values at low temperatures, an effect that we ascribe to the fact that the experimental T-jump responses likely contain a mixture of dissociation and hybridization events -- the latter of which persist to lower temperatures than the former -- whereas the computational analysis considers dissociation alone. Of the four sequences, GC-core shows the largest discrepancy between computation and experiment, although the general exponential trend is preserved. This may be a result of its high propensity to fray (cf.\ Section \ref{sec:GC-core}). In Fig.~\ref{fig:relaxation-comparison}b we expose a largely linear dependence of $k_d^\mathrm{fast}$ upon temperature for AT-all, GC-end, and GC-mix compared to an exponential dependence for GC-end. These trends can be understood in light of the comparatively larger enthalpic barrier for dissociation of the terminal G:C base pair in GC-end compared to that for the A:T terminal pair in the other three sequences. Again we see good agreement between the scaled computational predictions and the experimental T-jump measurements. The favorable comparison of computation and experiment provides support for the capacity of the 3SPN.2 model to reliably reproduce sequence-dependent trends in the slow and fast kinetics of the four DNA oligomers.

@@

We now proceeded to construct Markov state models (MSMs) from 1 ms of aggregated simulation trajectories for each of the four sequences at their respective melting temperatures to generate sequence-dependent kinetic models. MSMs define the long-lived metastable macrostates of the system, their equilibrium occupancies, and the equilibrium transition probabilities between them. As such, they are extremely valuable in providing both a quantitative predictive model and a physically comprehensible mechanistic understanding of the long-time dynamical evolution of the system between an ensemble of metastable macrostates. We present in Fig.~\ref{fig:allseq_table} the inferred MSMs for each of the four 10-base DNA sequences. Across all four sequences we identify a totality of seven metastable macrostates corresponding to the fully hybridized state (H) in which all native base pairings are intact, four shifted states in which the strands are translated out-of-register by two or four bases in the 5$^\prime$ (5S2, 5S4) or 3$^\prime$ (3S2, 3S4) direction, a frayed state (F4) -- unique to GC-core -- in which four terminal A:T base pairs are unbound, and the fully dissociated state (D). We present these seven macrostates in Fig.~\ref{fig:allseq_table}a along with schematic and cartoon renderings of representative microstates contained within each of these macrostates. In Fig.~\ref{fig:allseq_table}b we present the occupancy probabilities of each state at thermodynamic equilibrium. By virtue of the fact that each sequence is simulated at its corresponding melting temperature (T$_\mathrm{m}$), the probability of the dissociated state (D) is, by construction, approximately equal to the sum of the probabilities over the remaining six states (H, 5S2, 3S2, 5S4, 3S4, F). In Fig.~\ref{fig:allseq_table}c we present a visualization of the macrostate MSMs for each of the four sequences showing the connectivity between the identified macrostates. The macrostates are represented as orange circles in proportion to their equilibrium probabilities and the grey arrows indicate the probability of hopping from one macrostate to another under one time step of the kinetic model. The fluxes between the macrostates provide a wealth of high-level, interpretable information on the sequence-dependent metastable states and hydbrization/dehybridization pathways. Immediately, we identify that the AT-all sequence possesses a rich and complex dynamical landscape comprising six metastable states whereas at the other end of the spectrum GC-mix exhibits far simpler two-state ``all-or-nothing'' behavior. In Fig.~\ref{fig:allseq_table}d we present the so-called implied time scales of each MSM. These time scales correspond to the relaxation times of the DNA dimer among its constituent metastable macrostates. The leading implied time scale for each system corresponds to the characteristic time scale for hybridization/dehybridization. Since each system is simulated at the same concentration and at its respective melting temperature, it is not surprising that the leading time scale is approximately equal for all four systems and corresponds to the characteristic time scale for hybridization/dehybridization. The spectrum of higher order time scales corresponds to increasingly quicker relaxations between the metastable macrostates within the kinetic model and resolve the interesting sequence-dependent differences in the hybridization/dehybridization kinetic pathways. The total number of implied time scales is typically one fewer than the number of metastable macrostates and the existence of large implied time scales is indicative of slowly relaxing kinetic processes. The dense spectrum of slow implied time scales for AT-all is indicative of its relatively complex kinetic landscape whereas that for the two-state GC-mix comprises only a single time scale corresponding to hybridization/dehybridization. We now proceed to analyze in detail the sequence-dependent thermodynamics, kinetic, and mechanisms exposed by the four MSMs.

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{Fig2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
        \caption{Thermodynamic and kinetic predictions of the sequence-specific MSMs for AT-all, GC-end, GC-core, and GC-mix. (a) \textbf{Macrostates.} Schematic representation of the seven metastable macrostates occupied by one or more of the four sequences: fully hybridized state (H), two-base out-of-register shifted states in the 5$^\prime$ (5S2) or 3$^\prime$ (3S2) direction, four-base out-of-register shifted states in the 5$^\prime$ (5S4) or 3$^\prime$ (3S4) direction, and the fully dissociated state (D). The line drawings represent the 10-base self-complementary sequences, where red-to-blue contacts indicates (possible) WC base pairing and black indicates an unbound bases. Adjacent to each line drawing we provide representative molecular structures corresponding to that macrostate. (b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. (c) \textbf{Kinetic networks.} MSMs illustrating the kinetic network for each sequence. The orange circles correspond to the macrostates occupied by each sequence and are labeled by the macrostate codes reported in panel a. The radii of the circles is proportional to the logarithm of the equilibrium occupancy probabilities reported in panel b. Molecular renderings of an illustrative snapshot from the coarse-grained molecular simulations are provided next to each macrostate. The grey arrows between macrostates indicate the presence of a probability flux between this pair of states at equilibrium and the arrow thickness is proportional to the flux. (To avoid congesting the diagram, arrows are not reported for probability fluxes lower than $3\times10^{-6}$.) The numerical value overlaid on each arrow reports the conditional probability that a system occupying the macrostate at the start of the arrow at time $t$ will transition to the macrostate at the end of the arrow by time $(t+\tau)$, where $\tau$=1.2 ns is the lag time corresponding to a single time step of the MSM. Large orange circles correspond to thermodynamically favorable states and large grey arrows correspond to kinetically favorable transitions. (d) \textbf{Kinetic time scales.} Distribution of MSM implied time scales for each sequence. The leading implied time scale corresponds to the characteristic time scale for hybridization/dehybridization and is approximately equal for all systems since simulations were conducted at the same concentration and at the respective melting temperatures. The higher order implied time scales correspond to a spectrum of kinetic relaxations between the constituent macrostates in the MSM corresponding to shifted and/or frayed states.
    }
    \label{fig:allseq_table}
\end{figure}

@@

We first compare the thermodynamic predictions for the equilibrium macrostate probabilities to those of predictive sequence-dependent models for the thermodynamics of DNA association. The nearest neighbor (NN) model is a popular empirical model of DNA hybridization thermodynamics that predicts the free energy of duplex formation as a sum over helix initiation terms and the hybridization free energies of nearest neighbor pairs of bases that account for both the specific WC pairings and the modulating effects of the local (i.e., nearest neighbor) environment.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} The parameters of the NN model were estimated by regressing over 108 experimental measurements to furnish a predictive model for the free energy of association as a function of DNA sequence and explicitly account for stacking contributions of native pairs, internal mismatches, and dangling ends. We apply the NN model to predict the free energy $F^\mathrm{NN}$ of each of the macrostates occupied by each of the four sequences. A full accounting of our application of the NN model is provided in the \blauw{Supporting Information}. The free energy of each macrostate is related to its equilibrium occupancy probability $P$ via the statistical mechanical relationship $F = -k_B T \ln P + C$, where $T$ is temperature, $k_B$ is Boltzmann's constant, and $C$ is an additive constant reflecting our ignorance of the absolute scale of free energies. We use this relationship to convert the equilibrium occupancy probabilities predicted by our MSM and reported in Fig.~\ref{fig:allseq_table}b into free energies $F^\mathrm{MSM}$. The unknown additive constants preclude comparisons of absolute free energies between the MSM and NN model, but it is legitimate to compare relative free energies between pairs of macrostates since the additive constant cancels in taking differences. As such, we arbitrarily set the additive constant $C$ in both the MSM and NN model such that the hybridized state H defines a zero free energy reference state and we report the stability of all macrostates relative to the hybridized state as $\Delta F^\mathrm{NN} = F^\mathrm{NN} - F_H^\mathrm{NN}$ for the NN model and $\Delta F^\mathrm{MSM} = F^\mathrm{MSM} - F_H^\mathrm{MSM}$ for the MSM.

>>>>

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

In this work, we study a family of self-complementary pairs of 10-mer DNA oligomers using coarse-grained molecular simulation, machine learning of the slow dynamical modes, and data-driven inference of long-time kinetic models to establish new understanding of the influence of sequence upon hybridization/dehybridization kinetics and mechanisms. This family of oligomers -- 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and 5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix) -- was designed to probe the influence of the placement of a G:C pair \noteb{``of two G:C base pairs?''}within an otherwise repetitive A:T sequence and has been the subject of our prior experimental investigations using temperature-jump infrared spectroscopy and simple lattice models. \citep{Sanstead2016} We validate the new computational models of hybridization/dehybridization dynamics developed in this work against new experimental data and reinterpret our prior experimental observations in light of the new computational understanding. Consistent with previous studies, \citep{Hinckley2014Coarse-grainedEffects,Romano2013DNADependence,Araque2016LatticeCooperativity} we find the degree of repetitiveness in the sequence -- and therefore the kinetic accessibility and thermodynamic stability of out-of-register shifted states -- leads to richer dynamics populated by a diversity of long-lived metastable states. Our data-driven modeling and analysis rigorously quantifies these behaviors and furnishes accurate predictive models of the hybridization/dehybridization rates, dynamical pathways, and metastable states. Specifically, we demonstrate that disrupting repetitive stretches of A:T bases by placement of interrupting G:C base pairs enables us to tune the landscape from rich six-state to simple two-state ``all-or-nothing'' behavior, and the specific location of the interrupting pair can be used to modulate the stability of long-lived frayed states. Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to sculpt and qualitatively alter the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomena including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSimulations, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxations are, within a corrective scaling factor, in good agreement with observed experimental trends.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and slithering mechanisms.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values \noteb{I know we clarify this at the end of the paragraph, but maybe add ``our reported computational values in Fig.~\ref{fig:relaxation-comparison}'' to be clear?}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

@@

In Fig.~\ref{fig:relaxation-comparison}a we observe an exponential increase of $k_d^\mathrm{slow}$ with temperature, as expected from the large enthalpic barrier to duplex dissociation.\cite{Craig1971RelaxationOligonucleotides,Porschke1971CooperativeTransition,Williams1989LaserDGCATGC} Under the time and temperature calibration corrections, we see generally very good agreement between the computational and experimental curves. The computational rates do tend to systematically under-predict the experimental values at low temperatures, an effect that we ascribe to the fact that the experimental T-jump responses likely contain a mixture of dissociation and hybridization events -- the latter of which persist to lower temperatures than the former -- whereas the computational analysis considers dissociation alone \noteb{This effect is now minimized by the fact that we are using an experimental dissociation rates instead of the pure slow response (previous data showed worse agreement at low temp). Still, there could be some hybridization contribution so we could leave this sentence in}. Of the four sequences, GC-core shows the largest discrepancy between computation and experiment, although the general exponential trend is preserved. This may be a result of its high propensity to fray (cf.\ Section \ref{sec:GC-core}). In Fig.~\ref{fig:relaxation-comparison}b we expose a largely linear dependence of $k_d^\mathrm{fast}$ upon temperature for AT-all, GC-end, and GC-mix compared to an exponential dependence for GC-end. These trends can be understood in light of the comparatively larger enthalpic barrier for dissociation of the terminal G:C base pair in GC-end compared to that for the A:T terminal pair in the other three sequences. Again we see good agreement between the scaled computational predictions and the experimental T-jump measurements. The favorable comparison of computation and experiment provides support for the capacity of the 3SPN.2 model to reliably reproduce sequence-dependent trends in the slow and fast kinetics of the four DNA oligomers.

@@

We now proceeded to construct Markov state models (MSMs) from 1 ms of aggregated simulation trajectories for each of the four sequences at their respective melting temperatures to generate sequence-dependent kinetic models. MSMs define the long-lived metastable macrostates of the system, their equilibrium occupancies, and the equilibrium transition probabilities between them. As such, they are extremely valuable in providing both a quantitative predictive model and a physically comprehensible mechanistic understanding of the long-time dynamical evolution of the system between an ensemble of metastable macrostates. We present in Fig.~\ref{fig:allseq_table} the inferred MSMs for each of the four 10-base DNA sequences. Across all four sequences we identify a totality of seven metastable macrostates corresponding to the fully hybridized state (H) in which all native base pairings are intact, four shifted states in which the strands are translated out-of-register by two or four bases in the 5$^\prime$ (5S2, 5S4) or 3$^\prime$ (3S2, 3S4) direction, a frayed state (F4) -- unique to GC-core -- in which four terminal A:T base pairs are unbound, and the fully dissociated state (D). We present these seven macrostates in Fig.~\ref{fig:allseq_table}a along with schematic and cartoon renderings of representative microstates contained within each of these macrostates. In Fig.~\ref{fig:allseq_table}b we present the occupancy probabilities of each state at thermodynamic equilibrium. By virtue of the fact that each sequence is simulated at its corresponding melting temperature (T$_\mathrm{m}$), the probability of the dissociated state (D) is, by construction, approximately equal to the sum of the probabilities over the remaining six states (H, 5S2, 3S2, 5S4, 3S4, F). In Fig.~\ref{fig:allseq_table}c we present a visualization of the macrostate MSMs for each of the four sequences showing the connectivity between the identified macrostates. The macrostates are represented as orange circles in proportion to their equilibrium probabilities and the grey arrows indicate the probability of hopping from one macrostate to another under one time step of the kinetic model. The fluxes between the macrostates provide a wealth of high-level, interpretable information on the sequence-dependent metastable states and hybridization/dehybridization pathways. Immediately, we identify that the AT-all sequence possesses a rich and complex dynamical landscape comprising six metastable states whereas at the other end of the spectrum GC-mix exhibits far simpler two-state ``all-or-nothing'' behavior. In Fig.~\ref{fig:allseq_table}d we present the so-called implied time scales of each MSM. These time scales correspond to the relaxation times of the DNA dimer among its constituent metastable macrostates. The leading implied time scale for each system corresponds to the characteristic time scale for hybridization/dehybridization. Since each system is simulated at the same concentration and at its respective melting temperature, it is not surprising that the leading time scale is approximately equal for all four systems and corresponds to the characteristic time scale for hybridization/dehybridization. The spectrum of higher order time scales corresponds to increasingly quicker relaxations between the metastable macrostates within the kinetic model and resolve the interesting sequence-dependent differences in the hybridization/dehybridization kinetic pathways. The total number of implied time scales is typically one fewer than the number of metastable macrostates and the existence of large implied time scales is indicative of slowly relaxing kinetic processes. The dense spectrum of slow implied time scales for AT-all is indicative of its relatively complex kinetic landscape whereas that for the two-state GC-mix comprises only a single time scale corresponding to hybridization/dehybridization. We now proceed to analyze in detail the sequence-dependent thermodynamics, kinetic, and mechanisms exposed by the four MSMs.

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{Fig2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
        \caption{Thermodynamic and kinetic predictions of the sequence-specific MSMs for AT-all, GC-end, GC-core, and GC-mix. (a) \textbf{Macrostates.} Schematic representation of the seven metastable macrostates occupied by one or more of the four sequences: fully hybridized state (H), two-base out-of-register shifted states in the 5$^\prime$ (5S2) or 3$^\prime$ (3S2) direction, four-base out-of-register shifted states in the 5$^\prime$ (5S4) or 3$^\prime$ (3S4) direction, and the fully dissociated state (D). The line drawings represent the 10-base self-complementary sequences, where red-to-blue contacts indicates (possible) WC base pairing and black indicates an unbound bases. Adjacent to each line drawing we provide representative molecular structures corresponding to that macrostate. (b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. (c) \textbf{Kinetic networks.} MSMs illustrating the kinetic network for each sequence. The orange circles correspond to the macrostates occupied by each sequence and are labeled by the macrostate codes reported in panel a. The radii of the circles is proportional to the logarithm of the equilibrium occupancy populations reported in panel b. Molecular renderings of an illustrative snapshot from the coarse-grained molecular simulations are provided next to each macrostate. The grey arrows between macrostates indicate the presence of a probability flux between this pair of states at equilibrium and the arrow thickness is proportional to the flux. (To avoid congesting the diagram, arrows are not reported for probability fluxes lower than $3\times10^{-6}$.) The numerical value overlaid on each arrow reports the conditional probability that a system occupying the macrostate at the start of the arrow at time $t$ will transition to the macrostate at the end of the arrow by time $(t+\tau)$, where $\tau$=1.2 ns is the lag time corresponding to a single time step of the MSM. Large orange circles correspond to thermodynamically favorable states and large grey arrows correspond to kinetically favorable transitions. (d) \textbf{Kinetic time scales.} Distribution of MSM implied time scales for each sequence. The leading implied time scale corresponds to the characteristic time scale for hybridization/dehybridization and is approximately equal for all systems since simulations were conducted at the same concentration and at the respective melting temperatures. The higher order implied time scales correspond to a spectrum of kinetic relaxations between the constituent macrostates in the MSM corresponding to shifted and/or frayed states.
    }
    \label{fig:allseq_table}
\end{figure}

@@

We first compare the thermodynamic predictions for the equilibrium macrostate probabilities to those of predictive sequence-dependent models \noteb{does ``model'' imply we're looking at more than NN?} for the thermodynamics of DNA association. The nearest neighbor (NN) model is a popular empirical model of DNA hybridization thermodynamics that predicts the free energy of duplex formation as a sum over helix initiation terms and the hybridization free energies of nearest neighbor pairs of bases that account for both the specific WC pairings and the modulating effects of the local (i.e., nearest neighbor) environment.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} The parameters of the NN model were estimated by regressing over 108 experimental measurements to furnish a predictive model for the free energy of association as a function of DNA sequence and explicitly account for stacking contributions of native pairs, internal mismatches, and dangling ends. We apply the NN model to predict the free energy $F^\mathrm{NN}$ of each of the macrostates occupied by each of the four sequences. A full accounting of our application of the NN model is provided in the \blauw{Supporting Information}. The free energy of each macrostate is related to its equilibrium occupancy probability $P$ via the statistical mechanical relationship $F = -k_B T \ln P + C$, where $T$ is temperature, $k_B$ is Boltzmann's constant, and $C$ is an additive constant reflecting our ignorance of the absolute scale of free energies. We use this relationship to convert the equilibrium occupancy probabilities predicted by our MSM and reported in Fig.~\ref{fig:allseq_table}b into free energies $F^\mathrm{MSM}$. The unknown additive constants preclude comparisons of absolute free energies between the MSM and NN model, but it is legitimate to compare relative free energies between pairs of macrostates since the additive constant cancels in taking differences. As such, we arbitrarily set the additive constant $C$ in both the MSM and NN model such that the hybridized state H defines a zero free energy reference state and we report the stability of all macrostates relative to the hybridized state as $\Delta F^\mathrm{NN} = F^\mathrm{NN} - F_H^\mathrm{NN}$ for the NN model and $\Delta F^\mathrm{MSM} = F^\mathrm{MSM} - F_H^\mathrm{MSM}$ for the MSM.

As illustrated in Fig.~\ref{fig:NN_table}, we see that the MSM tends to predict higher relative free energies for the out-of-register shifted macrostates 5S2, 3S2, 5S4, and 3S4 compared to the NN model, such that the MSM predicts lower equilibrium occupancies of these states. The NN and MSM predictions for the frayed state F4 are in good agreement. Although the \noteb{shifted state} trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the predicted relative stability of the shifted states? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model, not least of which is the coarse-grained representation that integrates over atomic degrees of freedom, and in the NN model that was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end tend to destabilize the duplex.\citep{DiMichele2014EffectHybridization}
